---
title: "DataMining-Project"
output:
  pdf_document: default
  html_document: default
date: "2024-07-31"
---

## R Markdown

```{r}
rm(list = ls())
```

```{r}
suppressWarnings({ 
  library(readxl)
})
```

```{r}
suppressWarnings({
  df <- read_excel("khesarat_badaneh.xlsx")
  })
```

```{r}
summary(df)
```

```{r}
df$`مشکوک به تقلب` <- ifelse(!is.na(df$`مشکوک به تقلب`),1,0)
```

```{r}
table(df$`مشکوک به تقلب`)
str(df$`مشکوک به تقلب`)
```

```{r}
dim(df)
```

```{r}
suppressWarnings({
  library(dplyr)
})
```

```{r}
head(df)
```

```{r}
date_columns <- c("تاریخ انقضاء بیمه نامه","تاریخ شروع بیمه نامه"
                        ,"زمان ثبت","تاريخ صدور بيمه نامه","تاریخ تغییر وضعیت"
                        ,"تاریخ تشکیل پرونده"
                        ,"تاريخ صدور گواهينامه"
                        ,"تاریخ حادثه","تاریخ اعلام خسارت"
                        )
df <- df %>%
  mutate(across(all_of(date_columns), ~ as.Date(.x, format = "%Y/%m/%d")))
```

```{r}
numeric_columns <- c(
  "ردیف", "کد رایانه", "شماره پرونده", 
  "شماره سریال کروکی", "مبلغ بازیافت شده", "مبلع خسارت پرداخت شده", "مبلع کارشناسی پرداخت شده"
)
df <- df %>%
  mutate(across(all_of(numeric_columns), as.numeric))
```

### Remove Columns Which Contain Unusable Information

```{r}
df <- df[, !names(df) %in% c("کارشناس پرونده","شماره بیمه نامه", "کد رایانه","ردیف","شماره پرونده","زمان ثبت","شماره قرارداد")]
```

### Detecting Duplicate Values
```{r}
df_clean <- df[!duplicated(df),]
dim(df_clean)
```

```{r}
head(df_clean)

```

### Detecting Missing Values

```{r}
calculate_missing <- function(dataframe){
  y <- c()
  x <- c()
  n <- c()
  for (i in 1:ncol(dataframe)){
    n[i] <- sum(is.na(dataframe[,i]))
    if (n[i] != 0 & (n[i]/nrow(dataframe))>0.75){
      cat("the",i,"th column with missinig values percentages > 0.75",names(dataframe[,i]),"\n")
      cat("number of the ",i,"th column missinig values:",n[i], "\n")
      x <- c(x,i)
    }else if (n[i] != 0 & (n[i]/nrow(dataframe) < 0.75)){
      y <- c(y,i)
    }
  
  }
  result <- list(morethan_0.75 = x, lessthan_0.75 = y)
  return(result)
}
```

```{r}
m <- calculate_missing(df)
m
```

### Drop Features With More Than 75% Missing Values
```{r}
df_clean <- df[,-(m$morethan)]
head(df_clean)
```

```{r}
n <- calculate_missing(df_clean)
n
```

```{r}
str(df_clean[,n$lessthan])
```

### Impute Missing Value
```{r}
head(df_clean)

```

```{r}
numeric_columns_to_impute_df_clean <- c(49,42,48,47)
for (col in numeric_columns_to_impute_df_clean){
  df_clean[is.na(df_clean[,col]),col] <- mean(!is.na((df_clean[,col])), na.rm = TRUE)
  cat("number of missing values of col",names(df_clean[,col]),"after imputing:",sum(is.na(df_clean[,col])),"\n")
}

```

```{r}

categorical_columns_to_impute_df_clean <- c(7,12,13,17,19,46)


get_mode <- function(v) {
  uniq_v <- unique(v)
  uniq_v[which.max(tabulate(match(v, uniq_v)))]
}
for (col_index in categorical_columns_to_impute_df_clean) {
  col_name <- names(df_clean)[col_index]
  mode_value <- get_mode(na.omit(df_clean[[col_name]]))
  df_clean[[col_name]][is.na(df_clean[[col_name]])] <- mode_value
}
```

```{r}
fixed_value <- as.POSIXct("2024-01-01 00:00:00", format="%Y-%m-%d %H:%M:%S")
for (col in date_columns) {
  if (col %in% colnames(df_clean)) {
    df_clean[[col]][is.na(df_clean[[col]])] <- fixed_value
  } else {
    warning(paste("ستون", col, "در دیتافریم موجود نیست."))
  }
}

```

```{r}
for (col in 1:ncol(df_clean)){
  cat("number of missing values of col",names(df_clean[,col]),"after imputing:",sum(is.na(df_clean[,col])),"\n")
}
```

```{r}
head(df_clean)
```

### Encoded Character Columns

```{r}
categorical_columns <- c("آخرین شماره الحاقیه بیمه نامه
","شماره بیمه نامه", "مشکوک به تقلب", "واحد تشکیل دهنده پرونده", 
  "کارشناس پرونده", "خسارت خارج از مرز است ؟", "محل حادثه", 
  "نشاني دقيق محل حادثه", "علت حادثه", "نوع حادثه", "آيا تخريب صورت گرفته است ؟","گزارش", 
  "مقصر حادثه", "واحد تشکیل دهنده پرونده", "معرف بیمه نامه", "واحد صدور بیمه نامه", 
  "شماره قرارداد", "رشته", "وضعيت آخرين حواله"
)
df_clean <- df_clean %>%
  mutate(across(where(is.character), as.factor))
df_clean$`آخرین شماره الحاقیه بیمه نامه` <- as.factor(df_clean$`آخرین شماره الحاقیه بیمه نامه`)
```

```{r}
head(df_clean)
```

### Removing columns that have only one unique value(one level)

```{r}
single_level_factors <- sapply(df_clean, function(x) if (is.factor(x)) length(levels(x)) == 1 else FALSE)
df_clean <- df_clean[, !single_level_factors]
```

### Remove DateTime columns

```{r}
datetime_columns <- c("تاریخ انقضاء بیمه نامه","تاریخ شروع بیمه نامه"
                        ,"تاريخ صدور بيمه نامه","تاریخ تغییر وضعیت"
                        ,"تاریخ تشکیل پرونده"
                        ,"تاريخ صدور گواهينامه"
                        ,"تاریخ حادثه","تاریخ اعلام خسارت","ساعت حادثه"
                        )
df_clean <- df_clean[, !names(df_clean) %in% datetime_columns]

```


```{r}
df_clean_encoded <- df_clean
df_clean_encoded2 <- df_clean_encoded

```

```{r}
df_clean_encoded$target <- df_clean_encoded[["مشکوک به تقلب"]]
df_clean_encoded <- df_clean_encoded[, !names(df_clean_encoded) %in% c("مشکوک به تقلب")]
table(df_clean_encoded$target)
df_clean_encoded$target <- as.factor(df_clean_encoded$target)

```

```{r}
head(df_clean_encoded)
```

```{r}
str(df_clean_encoded)
```


```{r}
column_details <- lapply(df_clean_encoded, function(x) {
  paste("Type:", class(x), ", Number of Levels:", if (is.factor(x)) length(levels(x)) else "N/A")
})

column_info_detailed <- data.frame(
  Column_Name = names(column_details),
  Column_Details = unlist(column_details),
  stringsAsFactors = FALSE
)

print(column_info_detailed)

```

```{r}
suppressWarnings({
  library(car)
  library(caret)
})
```
### Balancing target column
```{r}
suppressWarnings({
  library(smotefamily)
})
```

```{r}
encoding_mappings <- list()

for (col in names(df_clean_encoded)) {
  if (is.factor(df_clean_encoded[[col]])) {
    original_levels <- levels(df_clean_encoded[[col]])
    encoding_mappings[[col]] <- setNames(seq_along(original_levels), original_levels)
    df_clean_encoded[[col]] <- as.numeric(df_clean_encoded[[col]])
  }
}
```

```{r}
factor_columns <- sapply(df_clean_encoded, is.factor)
original_levels <- lapply(df_clean_encoded[, factor_columns], levels)

df_clean_encoded[, factor_columns] <- lapply(df_clean_encoded[, factor_columns], function(x) as.numeric(factor(x, levels = levels(x))))

```


```{r}
smote_data <- SMOTE(df_clean_encoded[, -which(names(df_clean_encoded) == "target")], 
                    df_clean_encoded$target, K = 5, dup_size = 1000)

df_balanced <- data.frame(smote_data$data)
df_balanced$target <- as.factor(df_balanced$class)
df_balanced$class <- NULL
table(df_balanced$target)
```

```
decode_column <- function(encoded_col, original_levels) {
  factor_levels <- names(original_levels)
  decoded_col <- factor(encoded_col, levels = seq_along(factor_levels), labels = factor_levels)
  return(as.character(decoded_col))
}
colnames(df_balanced) = names(df_clean_encoded)
keep_as_double= c(33,34,35,36)
df_balanced[-keep_as_double] <- lapply(df_balanced[-keep_as_double], function(x) floor(x))
```

```
for (col in names(encoding_mappings)) {
    df_balanced[[col]] <- decode_column(df_balanced[[col]], encoding_mappings[[col]])
}

df_balanced <- df_balanced %>%
  mutate(across(where(is.character), as.factor))

head(df_balanced)
```

```{r}
head(df_balanced)
table(df_balanced$target)
```

```{r}
suppressWarnings({
  library(randomForest)
})
```

```{r}
suppressWarnings({
  library(xgboost)
})
```

```{r}
set.seed(123)
df_final <- df_balanced
df_final$target <- as.numeric(df_final$target)
df_final$target <- ifelse(df_final$target == 2, 1, 0)
```


```{r}
head(df_final)
table(df_final$target)
```

### Splitting data into train and test
```{r}
train_indices <- sample(seq_len(nrow(df_final)), size = 0.8 * nrow(df_final))
train_data <- df_final[train_indices, ]
test_data <- df_final[-train_indices, ]


train_x <- train_data %>% select(-target)
train_y <- train_data$target
test_x <- test_data %>% select(-target)
test_y <- test_data$target
```

```{r}
table(df_final$target)
```

### Feature Selection
```{r}
dtrain <- xgb.DMatrix(data = as.matrix(train_x), label = train_y)
dtest <- xgb.DMatrix(data = as.matrix(test_x), label = test_y)

params <- list(
  objective = "binary:logistic",
  eval_metric = "auc",
  max_depth = 2,
  eta = 0.05,
  nthread = 2,
  gamma = 1,
  subsample = 0.8,
  colsample_bytree = 0.8,
  lambda = 10,   # L2 regularization term 
  alpha = 10     # L1 regularization term
)

xgb_model <- xgb.train(params, dtrain, nrounds = 100, watchlist = list(val = dtest, train = dtrain), early_stopping_rounds = 10)

importance_matrix <- xgb.importance(feature_names = colnames(train_x), model = xgb_model)
print(importance_matrix)

top_features <- importance_matrix[importance_matrix$Gain > 0.05, ]$Feature
train_x_selected <- train_x %>% select(all_of(top_features))
test_x_selected <- test_x %>% select(all_of(top_features))
```
### Cross-Validation
```{r}
cv_folds <- 5
cv_model <- xgb.cv(
  params = params,
  data = dtrain,
  nrounds = 100,
  nfold = cv_folds,
  showsd = TRUE,
  stratified = TRUE,
  print_every_n = 10,
  early_stopping_rounds = 10,
  maximize = FALSE
)

best_nrounds <- cv_model$best_iteration
print(best_nrounds)
```

### Implement XGBOOST on data
```{r}

dtrain_selected <- xgb.DMatrix(data = as.matrix(train_x_selected), label = train_y)
dtest_selected <- xgb.DMatrix(data = as.matrix(test_x_selected), label = test_y)


xgb_model_selected <- xgb.train(params, dtrain_selected, nrounds = best_nrounds, watchlist = list(val = dtest_selected, train = dtrain_selected), early_stopping_rounds = 10)

preds <- predict(xgb_model_selected, dtest_selected)
pred_labels <- ifelse(preds > 0.5, 1, 0)
conf_matrix <- confusionMatrix(factor(pred_labels), factor(test_y))
print(conf_matrix)

```
